\documentclass[11pt]{article}
\usepackage{graphicx}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[shortlabels]{enumitem}
\usepackage[margin=1in]{geometry}

\newcommand{\R}{\mathbb{R}}

\newenvironment{solution}
  {\renewcommand\qedsymbol{$\blacksquare$}\begin{proof}[Solution]}
  {\end{proof}}

\setlength\parindent{0pt}

\begin{document}

	\hrule
	\begin{center}
        \textbf{MATH104: Topology}\hfill \textbf{Fall 2023}\newline

		{\Large Homework 5}

		David Yang
	\end{center}

\hrule

\vspace{1em}

\textit{Chapter 9 (The Fundamental Group) Problems.} \\

\underline{Section 53 (Covering Spaces), 53.3} \\

\textbf{Let $x_0$ and $x_1$ be points of the path-connected space $X$. Show that $\pi_1(X, x_0)$ is abelian if and only if every pair
$\alpha$ and $\beta$ of paths from $x_0$ to $x_1$, we have $\hat{\alpha} = \hat{b}$.}

\begin{solution}
We begin with the forward implication. Let $[f] \in \pi_1(X, x_0)$ and let $\alpha$ and $\beta$ be two paths from $x_0$ to $x_1$. 
Since $[\bar{\alpha}] * [\alpha] = e_{x_1}$ and $[\beta] * e_{x_1} * [\bar{\beta}] = e_{x_0}$, it follows  that
\[
    [f] = [\beta] * [\bar{\alpha}] * [\alpha] * [\bar{\beta}] * [f].
\]
Equivalently,
\[
    [f] = [\beta * \bar{\alpha}] * [\alpha * \bar{\beta}] * [f].
\]
Furthermore, note that $[\alpha * \bar{\beta}]$ is a loop based at $x_0$, so it is in $\pi_1(X, x_0)$, which is abelian. 
Consequently, $[\alpha * \bar{\beta}]$ commutes with $[f]$, so $[\alpha * \bar{\beta}] * [f] = [f] * [\alpha * \bar{\beta}]$. This gives us

\begin{align*}
    [f] &= [\beta * \bar{\alpha}] * [\alpha * \bar{\beta}] * [f] \\
    &= [\beta * \bar{\alpha}] * [f] * [\alpha * \bar{\beta}] \\
    &= [\beta] * [\bar{\alpha}] * [f] * [\alpha] * [\bar{\beta}]
\end{align*}
Finally, multiplying both sides by $[\bar{\beta}]$ on the left and by $[\beta]$ on the right and simplifying, we get that
\[
    [\bar{\beta}] * [f] * [\beta] = [\bar{\alpha}] * [f] * [\alpha]
\]
and this is equivalent to
\[
    \hat{\alpha}([f]) = \hat{\beta}([f]).
\]
We conclude that for any two paths $\alpha$ and $\beta$ from $x_0$ to $x_1$, $\hat{\alpha} = \hat{\beta}$, as desired. \\

It remains to show the reverse implication. Suppose that for any two paths $\alpha$ and $\beta$ of paths from $x_0$ to $x_1$, $\hat{\alpha} = \hat{b}$. Let
$[f_1]$ and $[f_2]$ be distinct path homotopy classes in $\pi_1(X, x_0)$. To show that $\pi_1(X, x_0)$ is abelian, we will show that $[f_1] * [f_2] = [f_2] * [f_1]$. \\

Note that $f_1 * \alpha$ and $f_2 * \alpha$ are two paths from $x_0$ to $x_1$, so we know that $\widehat{f_1 * \alpha} = \widehat{f_2 * \alpha}$. It follows that
$\widehat{f_1 * \alpha}([f_1]) = \widehat{f_2 * \alpha}([f_1])$, so
\[
    [\overline{f_1 * \alpha}] * [f_1] * [f_1 * \alpha] = [\overline{f_2 * \alpha}] * [f_1] * [f_2 * \alpha].
\]
Simplifying, we have that
\begin{align*}
    [\overline{f_1 * \alpha}] * [f_1] * [f_1 * \alpha] &= [\overline{f_2 * \alpha}] * [f_1] * [f_2 * \alpha] \\
    \implies [\bar{\alpha} * \bar{f_1}] * [f_1] * [f_1 * \alpha] &= [\bar{\alpha} * \bar{f_2}] * [f_1] * [f_2 * \alpha] \\
    \implies [\bar{\alpha}] * [\bar{f_1}] * [f_1] * [f_1] * [\alpha] &= [\bar{\alpha}] * [\bar{f_2}] * [f_1] * [f_2] * [\alpha].
\end{align*}
Multiplying both sides on the left by $[\alpha]$ and on the right by $[\bar{\alpha}]$ and simplifying further, we have that
\begin{align*}
    [\bar{\alpha}] * [\bar{f_1}] * [f_1] * [f_1] * [\alpha] &= [\bar{\alpha}] * [\bar{f_2}] * [f_1] * [f_2] * [\alpha]. \\
    \implies [f_1] &= [\bar{f_2}] * [f_1] * [f_2].
\end{align*}
Finally, multiplying both sides on the left by $[f_2]$, we get that 
\[
    [f_2] * [f_1] = [f_1] * [f_2]
\]
as desired. Thus, we conclude that for any two elements $[f_1]$ and $[f_2]$ in $\pi_1(X, x_0)$, $[f_1] * [f_2] = [f_2] * [f_1]$, so $\pi_1(X, x_0)$ is abelian, as desired.
\end{solution}

\newpage

\underline{Section 53 (Covering Spaces), 53.6(b)} \\

\textbf{Let $p\colon E \rightarrow B$ be a covering map.}

\begin{enumerate}[b)]
    \item \textbf{If $B$ is compact and $p^{-1}(b)$ is finite for each $b \in B$, then $E$ is compact.}
    
    \begin{solution}\footnote{constructed with Zoe's help. Also, I think there are a few steps that are not quite correct in the proof but I have tried for quite a while to fix them to no avail.
    Will be at Office Hours to talk about it.}
    Let $U = \bigcup\limits_{\alpha \in J} U_\alpha$ be an open cover of $E$. Since $p$ is a covering map, it is both an open map, so that $p(U_\alpha)$ is open, and surjective onto $B$. Consequently,
    $\bigcup\limits_{\alpha \in J} p(U_\alpha)$ is a union of open sets which covers $E$; equivalently, $\bigcup\limits_{\alpha \in J} p(U_\alpha)$ is an open cover of $B$. Let $b \in B$. Since $p$ is a covering map, 
    $b$ has a neighborhood $V_b$ that is openly covered by $p$. For each $b \in B$, pick an arbitrary $p(U_\alpha^b)$ containing it; we denote this as $\tilde{V}_\alpha^b$. It follows that the collection of
    \[
        \{ V_b \cap \tilde{V}_\alpha^b \mid b \in B \} 
    \]
    is an open covering of $B$ -- each element is the intersection of open sets which is open, and this set contains every point $b \in B$. 
    Since $B$ is compact, this open covering has a finite subcovering, which we will describe as
    \[
       \{ V_{b_i} \cap \tilde{V}_\alpha^{b_i} \mid i \in \{ 1, \dots, n \} \}
    \]
    for each $b_i \in B$. \\

    Consider now 
    \begin{align*}
        p^{-1} \left( \bigcup_{i \in \{ 1, \dots, n \}} V_{b_1} \cap \tilde{V}_\alpha^{b_1} \right) &= \bigcup_{i \in \{ 1, \dots, n \}} p^{-1}(V_{b_1}) \cap p^{-1}(\tilde{V}_\alpha^{b_1}) \\
        &= \bigcup_{i \in \{ 1, \dots, n \}} p^{-1}(V_{b_1}) \cap U_\alpha^{b_1}
    \end{align*}

    which is an open cover of $E$ -- each element is an intersection of open sets in $E$ and it covers $E$. Furthermore, it has a finite number of open sets, 
    as $p^{-1}(b)$ is finite for any $b \in B$.\footnote{I don't think is quite true as I'm working with an open set and using the fact that the inverse image of a point in $E$ is finite.} \\

    It follows that we can pick the set $U_\alpha^{b_i}$ for $i\in \{ 1, \dots, n \}$ to be an finite subcovering of the open cover $U$.\footnote{Also not sure about this, as I feel that I am neglecting the $p^{-1}(V_{b_1})$ part of the construction} Thus, any open cover of $E$ has a finite subcovering,
    and so $E$ is compact.
    \end{solution}
\end{enumerate}

\newpage

\textbf{Suppose $X$ is contractible. Prove that there is a bijection between $[X, Y]$ and the set of path components of $Y$.}

\begin{solution}\footnote{constructed with Phil's help.}
Since $X$ is contractible, the identity map $i_X$ is homotopic to a constant map. Let $F$ be this homotopy, the continuous map $F\colon X \times I \rightarrow  X$ satisfying
\[
    F(x, 0) = i_X(x) = x \text{ and } F(x, 1) = x_0
\]
for all $x \in X$ and some fixed $x_0 \in X$. \\

Let $f$ be a continuous map from $X$ into $Y$ where $f(x_0) = y_0$ for some $y_0 \in Y$. It follows that $f$ is homotopic to the constant map from $X$ to $Y$ sending any element in $X$ to $y_0$. 
The homotopy between these two maps, which we will denote as $G\colon X \times I \rightarrow Y$, is defined by $G(x, t) = f(F(x, t))$ and satisfies
\[
    G(x, 0) = f(F(x, 0)) = f(x) \text{ and } G(x, 1) = f(F(x, 1)) = f(x_0) = y_0.
\]
Furthermore, let $y_1 \in Y$ be in the same path component as $y_1$. We claim that $f$ is homotopic to the constant map from $X$ to $Y$ sending any element of $X$ to $y_0$. 
Since $y_0$ and $y_1$ are in the same path component of $Y$, there is some path $p$ connecting them. Let us parametrize $p$ so that $p\colon I \rightarrow Y$ satisfies $p(0) = y_0$ and $p(1) = y_1$. 
The homotopy between $f$ and the constant map from $X$ to $Y$ sending any element of $X$ to $y_1$ is $H\colon X \times I \rightarrow Y$ defined by
\[
    H(x, t) = \begin{cases}
        G(x, 2t) &\text{ if $t \leq \frac{1}{2}$}  \\
        p(2t - 1) &\text{ if $t > \frac{1}{2}$}
    \end{cases}.
\]
$H$ is continuous, as $H\left( x, \frac{1}{2} \right) = G(x, 1) = p(0) = y_0$; furthermore, 
\[
    H(x, 0) = G(x, 0) = f(x) \text{ and } H(x, 1) = p(1) = y_1.
\]

Thus, the maps $f$ of $X$ into $Y$ correspond directly to the constant maps from $X$ to $Y$. Furthermore, if two constant maps 
yield values in the same path component of $Y$, they are in the same homotopy class. Thus, we conclude that there is a bijection between $[X, Y]$, the set of
homotopy classes of maps of $X$ into $Y$ and the path components of $Y$, as desired.
\end{solution}

\end{document}
